---
title: "Appendix S3: BRMS models' validation after removing landscapes of the largest spatial extent"
output: html_document
---

This appendix presents a sensitivity analysis in which landscapes of the largest spatial extent (8 communities) were removed prior to model fitting to evaluate model robustness, so that all models were fitted to landscapes composed of 2–7 communities.

```{r General Settings, include=FALSE}
# include: whether to include anything from a code chunk in the output document --> when include = F, this whole chunk is excluded in the output!

# General Settings
knitr::opts_chunk$set(echo = F)
# Sets the R markdown options so that the code is displayed in the final html file. If it is set to FALSE then the code isn't displayed only the result is 

rm(list=ls()) #clear global environment

options(stringsAsFactors = F)
#to make sure that all strings are treated as characters and not factors unless specified

```

```{r Load Libraries and Data, include=FALSE}

# load libraries

library(tidyverse)  # tidy data for data manipulation and visualization
library(tidybayes)  # tidy data for Bayesian models
library(knitr)      # data manipulation
library(cowplot)    # for nice plot layout
library(brms)       # fitting Bayesian models
library(bayesplot)  # visualization of Bayesian models
library(rstan)      # bayesian models
library(cmdstanr)   # using the more-up-to-date cmdstanr instead of rstan as the underlying MCMC engine to avoid ***recursive gc invocation warnings
library(bayestestR) # helpful for some initial plotting 
library(see)        # the plots in the above package don't work without this 
library(loo)        # to do leave one out checks
library(MetBrewer)  # pretty colors

                  
# load custom functions
source("../../02_functions/CustomFunctions.R")

# load data
load("../../01_data/03_derived/scaling_data.RData")

```

```{r}
scaling_data<- filter(scaling_data, Area!= 8)
```

```{r}
theme_set(theme_tidybayes() + panel_border())
```

```{r remove outliers and normalize data}
filtered_data <- remove_outliers(data=scaling_data, vars=c("Spatial_AS","Gamma_stab"), quantile_cutoff = 0.997)
filtered_data <- normalize_min_max(data=filtered_data, var="Spatial_AS", epsilon = 0.0001, new_var_name = "Spatial_AS_normalized_offset")
```

# 1. Effects of Beta Diversity and Spatial Extent on Spatial Asynchrony

```{r import model 1}
model.SA<-readRDS("./models/SA_quantile.rds")
```


## 1.1 Model and convergence diagnostics 

**Table S3.1: Summary table of the BRMS model results.** The table summarizes the posterior distributions of the fixed effects and interaction terms. Reported values include the posterior means, standard errors, 95% credible intervals (CI), and effective sample sizes (Bulk ESS and Tail ESS) for each parameter.

```{r warning=FALSE}
print(model.SA, digits = 3)
```

Note the `Rhat` summary column: variation from 1.0 indicates the the model did not converge.

The Bayesian R-squared:
```{r}
bayes_R2(model.SA)
```

```{r message=FALSE}
mcmc_trace(as.array(model.SA)) + 
  scale_x_continuous(
    breaks = function(x) seq(min(x), max(x), length.out = 3),  # Show 3 ticks
    labels = function(x) sprintf("%.1f", x)  # Format ticks to 1 decimal place
  )
```

**Figure S3.1: Trace plots allow visually assessing chain convergence.**

## 1.2 Assessment of model fit and accuracy

### 1.2.1 Posterior predictive checks

We assess model fit using posterior predictive checks (PPC). PPCs compare the real data to the posterior distribution, conditioned on the observed data.

#### 1.2.1.1 Density plot

```{r}
pp_check(model.SA, ndraws = 500) + 
  theme_bw() + 
  xlab("Normalized Spatial Asynchrony") +
  ylab("Density")
```

**Figure S3.2: The density plot compares the real data (`y`, dark blue line) to the fitted draws of the models (`y_rep`, light blue lines).**

#### 1.2.1.2 Leave-One-Out Cross-Validation

```{r warning=FALSE, message=FALSE}
pp_check(model.SA, type="loo_pit_qq")+ 
  theme_bw() 
```

**Figure S3.3: The LOO-PIT (Leave-One-Out Probability Integral Transform) calibration plot compares the empirical cumulative distribution of the LOO-PIT values (solid blue line) against the uniform distribution (dashed diagonal line).**


### 1.2.2 Highest-density interval

We report 89% instead of 95% following common Bayesian practices, as it provides a better balance of precision and interpretability (McElreath, 2020).

```{r message=FALSE}
my.hdi.1 <- hdi(model.SA, ci = 0.89)
new_labels <- c( #create a vector with new parameter names that reflect the terminology of the ms
  "b_Beta_div" = "Beta Diversity",
  "b_Area" = "Spatial Extent",
  "b_Beta_div:Area" = "Beta Diversity x Spatial Extent"
)

  plot(my.hdi.1)+
  scale_y_discrete(labels = new_labels) # adjust labels

```

**Figure S3.4: We use the highest density interval (HDI) for the main effects in the model to characterize uncertainty of our posterior distributions.** 


## 1.3 The Hierarchical Effect of Species Richness on Spatial Asynchrony

**Table S3.2: Summary table of the hierarchical effects of planted species richness on spatial asynchrony.**

```{r}
ranef(model.SA)
```

```{r}
model.SA %>% spread_draws(r_div[div,]) %>% # Extracts the posterior draws for r_div by each level of div
  ungroup() %>%
  compare_levels(r_div, by = div) %>% # Compares posterior distributions of r_div across div levels
  ggplot(aes(y = div, x = r_div)) +
  stat_halfeye() +  # Adds a half-eye plot (density + credible intervals)
  geom_vline(xintercept = 0, linetype = "dashed") +  # Adds a dashed vertical reference line at x = 0
  # Axis labels
  xlab("Posterior Random Effect Value") +
  ylab("Levels of Planted Spcies\nRichness") +
  # Plot aesthetics
  theme(axis.title = element_text(size = 20, colour = "black"), #to modify axes appearance
        axis.line.x = element_line(colour = "black"),
        axis.line.y = element_line(colour = "black"),
        axis.text.x = element_text(size = 15,colour="black"),
        axis.text.y = element_text(size = 15,colour="black"),
        legend.text  = element_text(size = 15,colour="black"),
        legend.title = element_text(size = 20,colour="black"),
        legend.key.width = unit(1, "cm"),  # Increase legend key width to show linetypes clearly
        strip.text = element_text(size = 20),
        axis.ticks = element_line(colour="black"),
        panel.border = element_rect(colour="black", fill=NA))

```

**Figure S3.5. Posterior distributions of the random effects of planted species richness levels on gamma stability, comparing differences between species richness levels.** Black points represent posterior medians, and horizontal black lines indicate the 66% and 95% credible intervals. The dashed vertical line at x = 0 serves as a reference for no difference. Deviations from zero indicate the magnitude and direction of the differences in spatial asynchrony between species richness levels.


# 2. Effects of Spatial Asynchony and Spatial Extent on Gamma Stability

```{r import model 2}
model.GS<-readRDS("./models/GS_quantile.rds")
```

## 2.1 Model and convergence diagnostics 

**Table S3.3: Summary table of the BRMS model results.** The table summarizes the posterior distributions of the fixed effects and interaction terms. Reported values include the posterior means, standard errors, 95% credible intervals (CI), and effective sample sizes (Bulk ESS and Tail ESS) for each parameter.

```{r warning=FALSE}
print(model.GS, digits = 3)
```

Note the `Rhat` summary column: variation from 1.0 indicates the the model did not converge.

The Bayesian R-squared:
```{r}
bayes_R2(model.GS)
```

```{r message=FALSE}
mcmc_trace(as.array(model.GS)) + 
  scale_x_continuous(
    breaks = function(x) seq(min(x), max(x), length.out = 3),  # Show 5 ticks
    labels = function(x) sprintf("%.1f", x)  # Format ticks to 1 decimal place
  )
```

**Figure S3.6: Trace plots allow visually assessing chain convergence.**

## 2.2 Assessment of model fit and accuracy

### 2.2.1 Posterior predictive checks

We assess model fit using posterior predictive checks (PPC). PPCs compare the real data to the posterior distribution, conditioned on the observed data.

#### 2.2.1.1 Density plot

```{r}
pp_check(model.GS, ndraws = 500) + 
  theme_bw() + 
  xlab("Gamma Stability") +
  ylab("Density")
```

Figure S3.7: The density plot compares the real data (`y`, dark blue line) to the fitted draws of the models (`y_rep`, light blue lines).**

#### 2.2.1.2 Leave-One-Out Cross-Validation

```{r warning=FALSE, message=FALSE}
pp_check(model.GS, type="loo_pit_qq")+ 
  theme_bw() 
```

**Figure S3.8: The LOO-PIT (Leave-One-Out Probability Integral Transform) calibration plot compares the empirical cumulative distribution of the LOO-PIT values (solid blue line) against the uniform distribution (dashed diagonal line).**


### 2.2.2 Highest-density interval

We report 89% instead of 95% following common Bayesian practices, as it provides a better balance of precision and interpretability (McElreath, 2020).

```{r message=FALSE}
my.hdi.1 <- hdi(model.GS, ci = 0.89)
new_labels <- c( #create a vector with new parameter names that reflect the terminology of the ms
  "b_Spatial_AS" = "Spatial Asynchrony",
  "b_Area" = "Spatial Extent",
  "b_Spatial_AS:Area" = "Spatial Asynchrony x Spatial Extent"
)

  plot(my.hdi.1)+
  scale_y_discrete(labels = new_labels) # adjust labels

```

**Figure S3.9: We use the highest density interval (HDI) for the main effects in the model to characterize uncertainty of our posterior distributions.** 


## 2.3 The Hierarchical Effect of Species Richness on Spatial Asynchrony

**Table S3.4: Summary table of the hierarchical effects of planted species richness on gamma stability.**

```{r}
ranef(model.GS) 
```


```{r}
model.GS %>% spread_draws(r_div[div,]) %>% # Extracts the posterior draws for r_div by each level of div
  ungroup() %>%
  compare_levels(r_div, by = div) %>% # Compares posterior distributions of r_div across div levels
  ggplot(aes(y = div, x = r_div)) +
  stat_halfeye() +  # Adds a half-eye plot (density + credible intervals)
  geom_vline(xintercept = 0, linetype = "dashed") +  # Adds a dashed vertical reference line at x = 0
  # Axis labels
  xlab("Posterior Random Effect Value") +
  ylab("Levels of Planted Spcies\nRichness") +
  # Plot aesthetics
  theme(axis.title = element_text(size = 20, colour = "black"), #to modify axes appearance
        axis.line.x = element_line(colour = "black"),
        axis.line.y = element_line(colour = "black"),
        axis.text.x = element_text(size = 15,colour="black"),
        axis.text.y = element_text(size = 15,colour="black"),
        legend.text  = element_text(size = 15,colour="black"),
        legend.title = element_text(size = 20,colour="black"),
        legend.key.width = unit(1, "cm"),  # Increase legend key width to show linetypes clearly
        strip.text = element_text(size = 20),
        axis.ticks = element_line(colour="black"),
        panel.border = element_rect(colour="black", fill=NA))

```

**Figure S3.10. Posterior distributions of the random effects of planted species richness levels on gamma stability, comparing differences between species richness levels.** Black points represent posterior medians, and horizontal black lines indicate the 66% and 95% credible intervals. The dashed vertical line at x = 0 serves as a reference for no difference. Deviations from zero indicate the magnitude and direction of the differences in gamma stability between species richness levels.

